---
title: 'SaoMNK Tutorial: Decreasing density transitions'
author: "S. Downing"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

This project begins from the simple premise that the task of endogenizing the 
NK search landscape can be done by representing the search environment as a binary
bipartite network of `M` actors affiliating with `N` components. This bipartite 
network can be then be analyzed according to the Stochastic Actor-Oriented Model (SAOM) 
[(Snijders, 1996)](https://doi.org/10.1080/0022250X.1996.9990178). 

Thus, the Stochastic Actor-Oriented MNK model, abbrivated `SaoMNK`, is designed 
for running strategic search simulation, testing, and experimentation 
by leveraging the `RSiena` package, an R implementation of   
Simulation Investigation for Empirical Network Analysis (SIENA).

This tutorial offers a basic introduction to `SaoMNK`. 

```{r dependencies, include=FALSE}
rm(list=ls())
####
# ## General settings to hide code
# knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
knitr::opts_chunk$set(warning = FALSE)

# rm(list=ls())  ## uncomment to clear environ before run
########################
##
##   SAOM-NK Runscript:
##
##   1. Estimation Recovery 
##
##
#######################
library(plyr)
library(dplyr)
library(uuid)
library(ggpubr)
library(data.table)
library(DT)

#
.ts <- gsub('\\.','',as.character(as.numeric(Sys.time())))
#
# .uuid <- UUIDgenerate()
##  DIRECTORY
# grid.dir <- sprintf('_SAOMNK_sensitivity_start0_rand20__%s', .ts )
###############  Load R DEPENDENCIES ############################
## Directories
dir_ext <- 'D:\\Search_networks'
dir_r <- 'C:\\Users\\sdr8y\\OneDrive - University of Missouri\\Research\\Search_networks\\SaoMNK\\R'
###############  Load R6 Class DEPENDENCIES ############################
## Biparite Environment Search Simulation Class
SaomNkRSienaBiEnv <- source(file.path(dir_r, 'SAOM_NK_R6_model.R'))$value
```


```{r, warning=FALSE}
###############  Load R6 Class DEPENDENCIES ############################
## Biparite Environment Search Simulation Class
SaomNkRSienaBiEnv <- source(file.path(dir_r, 'SAOM_NK_R6_model.R'))$value
# ## RSiena search Class
# SaomNkRSienaBiEnv_search_rsiena <- source(file.path(dir_proj, 'SAOM_NK_R6_search_rsiena_model.R'))$value
## default settings: Users do not change; TODO: implment within restricted class attributes
DV_NAME <- 'self$bipartite_rsienaDV'
#
steps_per_actor <- 10
```



# 1. NO Endogenous Noise in Environment (Structural Effects = 0)

## 1.1 Environment Configuration

```{r}
#
environ_params <- list(
  M = 18,       ## Actors
  N = 9,       ## Components
  BI_PROB = 0, ## Environmental Density (DGP hyperparameter)
  component_matrix_start = 'rand', ##**TODO** Implement: 'rand','modular','semi-modular',...
  rand_seed = 1234,
  plot_init = F,
  name = '_test_tutorial_nb_'
)
#
env1 <- SaomNkRSienaBiEnv$new(environ_params)
```
```{r}
print(c(env1$M, env1$N))
```


```{r}
library(Matrix)  # For block diagonal matrices
#
create_block_diag <- function(N, B) {
  # Determine the approximate block size
  block_sizes <- rep(N %/% B, B)  # Equal-sized blocks
  block_sizes[1:(N %% B)] <- if (N %% 2 == 0) {
    block_sizes[1:(N %% B)]   # Adjust for remainder
  } else {
    block_sizes[1:(N %% B)] + 1
  }
  # Create individual binary blocks
  blocks <- lapply(block_sizes, function(s) matrix(1, nrow = s, ncol = s))
  # Combine blocks into a block diagonal matrix
  block_diag_matrix <- as.matrix(bdiag(blocks))  # Convert sparse to standard matrix
  # Ensure exact NxN dimensions
  block_diag_matrix <- block_diag_matrix[1:N, 1:N]  # Trim to N x N
  # return
  return(block_diag_matrix)
}
# Example usage:
N <- 6  # Number of rows/columns
B <- 2  # Number of blocks
block_matrix <- create_block_diag(8, 2)
# Print the matrix
print(block_matrix)
```


## 1.2 Structrual Model

SAOM Objective Function serves as the stochastic actor's utility function for strategic search. 

```{r}
#
strategies <- list(
  egoX   =   c(-1, 0, 1),
  inPopX =   c( 1, 0, -1)
)

## 2.b. Component Payoffs vector
set.seed(12345)
component_payoffs <-  runif(environ_params$N, min = 0, max = 1)
## 2. Strategies sets the objective function as a linear combination of network stats across DVs
#
actor_strats_list <- lapply(strategies, function(strat) rep(strat,  environ_params$M/length(strat)) )
#
structure_model <- list(
  dv_bipartite = list(
    name = 'self$bipartite_rsienaDV',
    effects = list( ##**STRUCTURAL EFFECTS -- dyadic/network endogeneity sources**
      list(effect='density', parameter= 0, dv_name=DV_NAME, fix=T ), ##interaction1 = NULL
      list(effect='inPop',   parameter= 0, dv_name=DV_NAME, fix=T ), #interaction1 = NUL
      list(effect='outAct',  parameter= 0, dv_name=DV_NAME, fix=T )
    ),
    ## COVARIATE EFFECTS
    coCovars = list( 
      ##** COMPONENTS : MONADIC CONSTANT COVARIATE EFFECTS **##
      list(effect='altX',   parameter= 1, dv_name=DV_NAME, fix=T,
           interaction1='self$component_1_coCovar', x = component_payoffs 
      ),
      ##** STRATEGIES : MONADIC CONSTANT COVARIATE EFFECTS **##
      list(effect='egoX',   parameter= .1,  dv_name=DV_NAME, fix=T,
           interaction1='self$strat_1_coCovar',   x = actor_strats_list[[1]] 
      ), #interaction1 = NULL
      list(effect='inPopX', parameter= .1,  dv_name=DV_NAME, fix=T,
           interaction1='self$strat_2_coCovar',  x = actor_strats_list[[2]] 
      )
    ),
    ##**MONADIC TIME-VARYING COVARIATE EFFECTS -- DYNAMIC STRATEGY PROGRAMS**
    varCovars = list(),
    ##**DYADIC CONSTANT COVARIATE EFFECTS -- EXOGENOUS INTERACTION MATRIX**
    coDyadCovars = list(
       list(effect='XWX',   parameter= 1, dv_name=DV_NAME, fix=T,
           interaction1='self$component_1_coDyadCovar', 
           x = create_block_diag(environ_params$N, round(environ_params$N/3))
      )
    ),
     ##**DYADIC TIME-VARYING COVARIATE EFFECTS -- DYNAMIC INTERACTION MATRIX**
    varDyadCovars = list(),
    interactions = list(
      list(effect='altX|XWX',   parameter= 1, dv_name=DV_NAME, fix=T,
       interaction1='self$component_1_coCovar', 
       interaction2='self$component_1_coDyadCovar'
      )
    )
  )
)
```



```{r}
env1$preview_effects(structure_model, filter=FALSE)
```



```{r}
# ## Uncomment for HTML output with filterable data table
# env1$preview_effects(structure_model, filter=TRUE)
```



## 1.3 Run RSiena Search Process

```{r}

## TODO:  PICK UP WITH coDydCovar Interation Matrix

## Run Rsiena search using variable parameters in theta_matrix
env1$search_rsiena(
  structure_model,
  iterations = env1$M * steps_per_actor,
  digits = 4,
  run_seed = 12345
)
```






```{r}
## 1st and last state of the bipartite matrix system
print(env1$bipartite_matrix_init )
print(env1$bi_env_arr[,, dim(env1$bi_env_arr)[3] ] )
```

## 1.4 Print Bipartite Search Network Snapshots 

Snapshopts of the the biparite network and the social and component epistasis interactions
are taken at fixed step intervals to show the evolving multidimensional coupled search 
environment (actors and components) 


```{r, fig.height=5, fig.width=11}
# snapshot_ids <- round( seq(1, dim(env1$bi_env_arr)[3], length=3 ) )
snapshot_ids <- c(1, 1+ (1:3)*env1$M)
for (i in 1:length(snapshot_ids)) {
  step <- snapshot_ids[ i ]
  mat  <- env1$bi_env_arr[,,step]
  env1$plot_bipartite_system_from_mat(mat, step)
}
```



## 1.5 Plot Actor Degrees (Component Scope and Common Affiliation Social Ties)


Time series is simulated decision steps.

```{r, fig.height=6, fig.width=6.5, fig.dpi=800}
##
env1$plot_actor_degrees(loess_span = 0.35)
```



## 1.6 Plot Component Degrees (Membership/Popularity and Common Actor Epistasis Ties)

```{r, fig.height=6, fig.width=6.5, fig.dpi=800}
##
env1$plot_component_degrees(loess_span = 0.35)

```




## 1.7 Plot Utility by Strategy over Time (Simulated Decision Steps) 

```{r, warning=FALSE, echo=FALSE}

env1$plot_strategy_utility() + 
  ggtitle('Utility by Strategy\nLoess Curves vs. Periodwise Avg. in Black')

```



## 1.8 Plot Utility by Actor over Time (Simulated Decision Steps) 

```{r, warning=FALSE, echo=FALSE}

env1$plot_actor_utility() + 
  ggtitle('Utility by Actor\nLoess Curves vs. Periodwise Avg. in Black')

```



## 1.9 Plot Contributions to Actor Utility colored by Strategy

Time series is simulated decision steps.

```{r,  fig.height=11, fig.width=8, fig.dpi=800}
##
env1$plot_utility_components(loess_span=0.35)
```




## 1.10 Plot Actor Utility Timeseries and Distribution by Strategy

```{r, fig.height=6, fig.width=8, fig.dpi=800}
env1$plot_actor_utility_strategy_summary(loess_span=0.5)
```





## 1.11  Fitness Landscape Peaks Distributions 

Number of changes to component ties as distance from counterfactual affiliation configurations

```{r}
## Use random uniform values for all combinations like traditional NK model
env1$compute_fitness_landscape(component_coCovar=NA)
```


```{r}
## Use the component values as the means of random noise for different payoff combinations 
env1$compute_fitness_landscape(component_coCovar=1)
```



```{r}
## LOWER SD of component values noise for different payoff combinations 
env1$compute_fitness_landscape(component_coCovar=1, component_value_sd=0.01)
```


```{r}
## HIGHER SD of component values noise for different payoff combinations 
env1$compute_fitness_landscape(component_coCovar=1, component_value_sd=0.5)
```


















# 2. Environmental WITH Endogenous Noise 

Structural Effects in range [-1, 0.4]

## 2.1 Environmental Configuration 

Reuse existing `environ_config`

## 2.2 Structrual Model

SAOM Objective Function serves as the stochastic actor's utility function for strategic search. 

```{r}
#
strategies <- list(
  egoX   =   c(-1, 0, 1),
  inPopX =   c( 1, 0, -1)
)

## 2.b. Component Payoffs vector
set.seed(12345)
component_payoffs <-  runif(environ_params$N, min = 0, max = 1)
## 2. Strategies sets the objective function as a linear combination of network stats across DVs
#
actor_strats_list <- lapply(strategies, function(strat) rep(strat,  environ_params$M/length(strat)) )
#
nsteps <- env1$M * steps_per_actor ## keep from above
#
structure_model <- list(
  dv_bipartite = list(
    name = 'self$bipartite_rsienaDV',
    effects = list( ##**STRUCTURAL EFFECTS -- dyadic/network endogeneity sources**
      list(effect='density', parameter= -.5, fix=T, dv_name=DV_NAME), ##interaction1 = NULL
      list(effect='inPop',   parameter= .1, fix=T, dv_name=DV_NAME), #interaction1 = NUL
      list(effect='outAct',  parameter= .1, fix=T, dv_name=DV_NAME)
    ),
    ## COVARIATE EFFECTS
    coCovars = list( 
      ##** COMPONENTS : MONADIC CONSTANT COVARIATE EFFECTS **##
      list(effect='altX',   parameter= 1, dv_name=DV_NAME, fix=T,
           interaction1='self$component_1_coCovar', x = component_payoffs 
      ),
      ##** STRATEGIES : MONADIC CONSTANT COVARIATE EFFECTS **##
      list(effect='egoX',   parameter= .1,  dv_name=DV_NAME, fix=T,
           interaction1='self$strat_1_coCovar',   x = actor_strats_list[[1]] 
      ), #interaction1 = NULL
      list(effect='inPopX', parameter= .1,  dv_name=DV_NAME, fix=T,
           interaction1='self$strat_2_coCovar',  x = actor_strats_list[[2]] 
      )
    ),
    ##**MONADIC TIME-VARYING COVARIATE EFFECTS -- DYNAMIC STRATEGY PROGRAMS**
    varCovars = list(),
    ##**DYADIC CONSTANT COVARIATE EFFECTS -- EXOGENOUS INTERACTION MATRIX**
    coDyadCovars = list(
       list(effect='XWX',   parameter= 1, dv_name=DV_NAME, fix=T,
           interaction1='self$component_1_coDyadCovar', 
           x = create_block_diag(environ_params$N, 2)
      )
    ),
     ##**DYADIC TIME-VARYING COVARIATE EFFECTS -- DYNAMIC INTERACTION MATRIX**
    varDyadCovars = list(),
    interactions = list()
  )
)


```



## 2.3 Run RSiena Search Process

```{r}
## Run Rsiena search using variable parameters in theta_matrix
env1$search_rsiena(
  structure_model,
  iterations = env1$M * steps_per_actor,
  digits = 4,
  run_seed = 12345
)

```



## 2.4 Print Bipartite Search Network Snapshots 

Snapshopts of the the biparite network and the social and component epistasis interactions
are taken at fixed step intervals to show the evolving multidimensional coupled search 
environment (actors and components) 


```{r, fig.height=5, fig.width=11}
# snapshot_ids <- round( seq(1, dim(env1$bi_env_arr)[3], length=3 ) )
snapshot_ids <- c(1, 1+ (1:3)*env1$M)
for (i in 1:length(snapshot_ids)) {
  step <- snapshot_ids[ i ]
  mat <- env1$bi_env_arr[,,step]
  env1$plot_bipartite_system_from_mat(mat, step)
}
```



## 2.5 Plot Actor Degrees (Component Scope and Common Affiliation Social Ties)


Time series is simulated decision steps.

```{r, fig.height=6, fig.width=6.5, fig.dpi=800}
##
env1$plot_actor_degrees(loess_span = 0.25)
```



## 2.6 Plot Component Degrees (Membership/Popularity and Common Actor Epistasis Ties)

```{r, fig.height=6, fig.width=6.5, fig.dpi=800}
##
env1$plot_component_degrees(loess_span = 0.35)

```




## 2.7 Plot Utility by Strategy over Time (Simulated Decision Steps) 

```{r, warning=FALSE, echo=FALSE}

env1$plot_strategy_utility() + 
  ggtitle('Utility by Strategy\nLoess Curves vs. Periodwise Avg. in Black')

```



## 2.8 Plot Utility by Actor over Time (Simulated Decision Steps) 

```{r, warning=FALSE, echo=FALSE}

env1$plot_actor_utility() + 
  ggtitle('Utility by Actor\nLoess Curves vs. Periodwise Avg. in Black')

```



## 2.9 Plot Contributions to Actor Utility colored by Strategy

Time series is simulated decision steps.

```{r,  fig.height=10, fig.width=8, fig.dpi=800}
##
env1$plot_utility_components(loess_span=0.35)
```




## 2.10 Plot Actor Utility Timeseries and Distribution by Strategy

```{r, fig.height=6, fig.width=8, fig.dpi=800}
env1$plot_actor_utility_strategy_summary()
```





## 2.11  Fitness Landscape Peaks Distribution  

Number of changes to component ties as distance from counterfactual affiliation configurations


```{r}
## Use random uniform values for all combinations like traditional NK model
env1$compute_fitness_landscape(component_coCovar=NA)
```


```{r}
## Use the component values as the means of random noise for different payoff combinations 
env1$compute_fitness_landscape(component_coCovar=1)
```



```{r}
## LOWER SD of component values noise for different payoff combinations 
env1$compute_fitness_landscape(component_coCovar=1, component_value_sd=0.01)
```


```{r}
## HIGHER SD of component values noise for different payoff combinations 
env1$compute_fitness_landscape(component_coCovar=1, component_value_sd=0.5)
```

```{r}
## 1st and last state of the bipartite matrix system
# dim(env1$fitness_landscape)

hist(env1$fitness_landscape[ 1, , dim(env1$fitness_landscape)[3]-1 ],
     main='Fitness Distribution of Landscape Draw 1', xlab='Fitness')
```

```{r}

hist(env1$fitness_landscape[ 1, , dim(env1$fitness_landscape)[3] ],
     main='Peaks Distribution over configurations in Landscape Draw 1', xlab='Peak')
```


```{r}
env1$fitness_landscape[1,,]



```





```{r, fig.height=8, fig.width=10, fig.dpi=800}
    
self <- env1

theta <- self$rsiena_model$thetaUsed

# step_ids <- round(c(1,8,64,512, self$rsiena_model$n3/2, self$rsiena_model$n3))
acpds <- seq(1, nrow(theta), by= self$M )
step_ids <- c(acpds[1:min(3, length(acpds))])
#
step_id <- step_ids[ 2 ]

  # step_id <- 1
  cat(sprintf('\nstep %s: actors ', step_id))
  
  bi_env_mat_step <- self$bi_env_arr[, , step_id]
  
  act_counterfacts <- list()
  
  i <- 1
  
  # for (i in 1:min(3, self$M)) {
    cat(sprintf(' %s ', i))
    
    
    ##**ACTOR i DECISION PERSPECTIVE** 0000000000000000000000000000000000
    
    ##**TODO**
    ## ALL actor-component counterfactual configuations for Actor i  (2^N rows by N cols)
    iland <- expand.grid(lapply(1:self$N, function(x) 0:1 ))
    iland_config_step_row_id <- which(apply(iland, 1, function(x) all(x == bi_env_mat_step[i,]) ))
    
    fit_vec <- env1$fitness_landscape[1, , 1 + 2*env1$N]
    
    ## fitness
    fitness_land1 <- env1$fitness_landscape[1, , 1:env1$N]
    #
    if ( ! all( iland == fitness_land1) ) {
      stop('fitness landscape configurations do not match iland configurations')
    }
    
    
    
    ##**TODO**
    ##**TODO** COMBINE utility and 'fitness' measures in dataframe by configuration with distance from actor's current configuration
    ##*##**TODO**
    ##**TODO**
    ##**TODO  CHANGE TO ACTOR-SPECIFIC LANDSCAPES **
    tmpmat <- bi_env_mat_step
    ## ifit dimensions [ M, 2^N ]
    iutil <- apply(iland, 1, function(x){
      tmpmat[i,] <- x  ## set counterfactual actor-component configuration
      config_util <- if('matrix' %in% class(theta)) {
        self$get_struct_mod_stats_mat_from_bi_mat( tmpmat ) %*% theta[step_id,] ## variable thetas
      } else {
        self$get_struct_mod_stats_mat_from_bi_mat( tmpmat ) %*% theta   ## fixed thetas
      }
      return( config_util ) ## compute utility vector
    }) 
    
   
    # plot(ifit[actor_i, ])
    
    # apply
    
    # ids.max <- which(ifit == max(ifit), arr.ind = TRUE) 
    # fits.max <- ifit[ ids.max ]
    # nmax <- length(fits.max)
    
    # ifit
    
    # ids.max[1,]
    
    # act_counterfacts[[i]] <- list(
    #   ifit = ifit, #Given all other ties, Actor i's configurations applied to utility func for all actors
    #   ids.max = ids.max,
    #   fits.max = fits.max,
    #   nmax = nmax
    # )
    
    ## distances of each counterfactual configuration 
    actor_i <- i
    start_actor_config <- bi_env_mat_step[actor_i, ]
    start_actor_config_str <- paste(start_actor_config, collapse = '|')
    dist_counterfac <- iland - matrix( rep(start_actor_config, 2^self$N), byrow=T, ncol=self$N)
    z <- t(sapply( 1:nrow(dist_counterfac),  function(config_i) {
        x <- dist_counterfac[config_i,  ]   ## dist_counterfac[MxN]
        util <- iutil[actor_i , config_i]  ## iutil[MxN]
        fit <- fit_vec[config_i]
        c(dist=sum(x!=0),
          drop=sum(x==-1), 
          nochange=sum(x==0),
          add=sum(x==1), 
          config_str = paste( iland[config_i, ], collapse = '|'),
          config_dist_str=paste(x, collapse = '|'), 
          fitness=fit, 
          utility=util
          )
      })) %>% tidyr::as_tibble() %>% 
      mutate(start_str=start_actor_config_str, 
             chain_step_id=step_id,
             fitness = as.numeric(fitness),
             utility = as.numeric(utility),
             dist = as.numeric(dist),
             drop = as.numeric(drop),
             nochange = as.numeric(nochange),
             add = as.numeric(add)
             )


plt <- z %>% ggplot(aes(x=utility, y=fitness, color=factor(add))) +
  geom_smooth(alpha=.1, method='lm') +
  geom_point(alpha=.2) +
  facet_grid(factor(dist) ~ . , scales='free_y') +
  geom_vline(aes(xintercept = mean, color=factor(dist)), data=z%>%group_by(dist,chain_step_id)%>%summarize(mean=mean(utility))) +
# facet_grid(.~dist) +
  theme_bw() + 
  ggtitle(sprintf('Actor %s Fitness as function of Utility\nby configuration distance (number of tie changes from current config.)', actor_i))

print(plt)

(fita <- lm(fitness ~ utility*factor(add), data=z ))
print(summary(fita))
(fitd <- lm(fitness ~ utility*factor(drop), data=z ))
print(summary(fitd))
    
```


```{r, fig.height=8, fig.width=10, fig.dpi=800}
   self <- env1

theta <- self$rsiena_model$thetaUsed

# step_ids <- round(c(1,8,64,512, self$rsiena_model$n3/2, self$rsiena_model$n3))
acpds <- seq(1, nrow(theta), by= self$M )
step_ids <- c(acpds[1:min(3, length(acpds))])
#
step_id <- step_ids[ 2 ]

  # step_id <- 1
  cat(sprintf('\nstep %s: actors ', step_id))
  
  bi_env_mat_step <- self$bi_env_arr[, , step_id]
  
  act_counterfacts <- list()
  
  i <- 2
  
  # for (i in 1:min(3, self$M)) {
    cat(sprintf(' %s ', i))
    
    
    ##**ACTOR i DECISION PERSPECTIVE** 0000000000000000000000000000000000
    
    ##**TODO**
    ## ALL actor-component counterfactual configuations for Actor i  (2^N rows by N cols)
    iland <- expand.grid(lapply(1:self$N, function(x) 0:1 ))
    iland_config_step_row_id <- which(apply(iland, 1, function(x) all(x == bi_env_mat_step[i,]) ))
    
    fit_vec <- env1$fitness_landscape[1, , 1 + 2*env1$N]
    
    ## fitness
    fitness_land1 <- env1$fitness_landscape[1, , 1:env1$N]
    #
    if ( ! all( iland == fitness_land1) ) {
      stop('fitness landscape configurations do not match iland configurations')
    }
    
    
    
    ##**TODO**
    ##**TODO** COMBINE utility and 'fitness' measures in dataframe by configuration with distance from actor's current configuration
    ##*##**TODO**
    ##**TODO**
    ##**TODO  CHANGE TO ACTOR-SPECIFIC LANDSCAPES **
    tmpmat <- bi_env_mat_step
    ## ifit dimensions [ M, 2^N ]
    iutil <- apply(iland, 1, function(x){
      tmpmat[i,] <- x  ## set counterfactual actor-component configuration
      config_util <- if('matrix' %in% class(theta)) {
        self$get_struct_mod_stats_mat_from_bi_mat( tmpmat ) %*% theta[step_id,] ## variable thetas
      } else {
        self$get_struct_mod_stats_mat_from_bi_mat( tmpmat ) %*% theta   ## fixed thetas
      }
      return( config_util ) ## compute utility vector
    }) 
    
   
    # plot(ifit[actor_i, ])
    
    # apply
    
    # ids.max <- which(ifit == max(ifit), arr.ind = TRUE) 
    # fits.max <- ifit[ ids.max ]
    # nmax <- length(fits.max)
    
    # ifit
    
    # ids.max[1,]
    
    # act_counterfacts[[i]] <- list(
    #   ifit = ifit, #Given all other ties, Actor i's configurations applied to utility func for all actors
    #   ids.max = ids.max,
    #   fits.max = fits.max,
    #   nmax = nmax
    # )
    
    ## distances of each counterfactual configuration 
    actor_i <- i
    start_actor_config <- bi_env_mat_step[actor_i, ]
    start_actor_config_str <- paste(start_actor_config, collapse = '|')
    dist_counterfac <- iland - matrix( rep(start_actor_config, 2^self$N), byrow=T, ncol=self$N)
    z <- t(sapply( 1:nrow(dist_counterfac),  function(config_i) {
        x <- dist_counterfac[config_i,  ]   ## dist_counterfac[MxN]
        util <- iutil[actor_i , config_i]  ## iutil[MxN]
        fit <- fit_vec[config_i]
        c(dist=sum(x!=0),
          drop=sum(x==-1), 
          nochange=sum(x==0),
          add=sum(x==1), 
          config_str = paste( iland[config_i, ], collapse = '|'),
          config_dist_str=paste(x, collapse = '|'), 
          fitness=fit, 
          utility=util
          )
      })) %>% tidyr::as_tibble() %>% 
      mutate(start_str=start_actor_config_str, 
             chain_step_id=step_id,
             fitness = as.numeric(fitness),
             utility = as.numeric(utility),
             dist = as.numeric(dist),
             drop = as.numeric(drop),
             nochange = as.numeric(nochange),
             add = as.numeric(add)
             ) 

plt <- z %>% ggplot(aes(x=utility, y=fitness, color=factor(add))) + 
  geom_smooth(alpha=.1, method='lm') +
  geom_point(alpha=.2) + 
  facet_grid(factor(dist) ~ . , scales='free_y') +
  geom_vline(aes(xintercept = mean, color=factor(dist)), data=z%>%group_by(dist,chain_step_id)%>%summarize(mean=mean(utility))) +
# facet_grid(.~dist) +
  theme_bw() + 
  ggtitle(sprintf('Actor %s Fitness as function of Utility\nby configuration distance (number of tie changes from current config.)', actor_i))

print(plt)
    

(fita <- lm(fitness ~ utility*factor(add), data=z ))
print(summary(fita))
(fitd <- lm(fitness ~ utility*factor(drop), data=z ))
print(summary(fitd))

```

```{r, fig.height=8, fig.width=10, fig.dpi=800}
    self <- env1

theta <- self$rsiena_model$thetaUsed

# step_ids <- round(c(1,8,64,512, self$rsiena_model$n3/2, self$rsiena_model$n3))
acpds <- seq(1, nrow(theta), by= self$M )
step_ids <- c(acpds[1:min(3, length(acpds))])
#
step_id <- step_ids[ 2 ]

  # step_id <- 1
  cat(sprintf('\nstep %s: actors ', step_id))
  
  bi_env_mat_step <- self$bi_env_arr[, , step_id]
  
  act_counterfacts <- list()
  
  i <- 3
  
  # for (i in 1:min(3, self$M)) {
    cat(sprintf(' %s ', i))
    
    
    ##**ACTOR i DECISION PERSPECTIVE** 0000000000000000000000000000000000
    
    ##**TODO**
    ## ALL actor-component counterfactual configuations for Actor i  (2^N rows by N cols)
    iland <- expand.grid(lapply(1:self$N, function(x) 0:1 ))
    iland_config_step_row_id <- which(apply(iland, 1, function(x) all(x == bi_env_mat_step[i,]) ))
    
    fit_vec <- env1$fitness_landscape[1, , 1 + 2*env1$N]
    
    ## fitness
    fitness_land1 <- env1$fitness_landscape[1, , 1:env1$N]
    #
    if ( ! all( iland == fitness_land1) ) {
      stop('fitness landscape configurations do not match iland configurations')
    }
    
    
    
    ##**TODO**
    ##**TODO** COMBINE utility and 'fitness' measures in dataframe by configuration with distance from actor's current configuration
    ##*##**TODO**
    ##**TODO**
    ##**TODO  CHANGE TO ACTOR-SPECIFIC LANDSCAPES **
    tmpmat <- bi_env_mat_step
    ## ifit dimensions [ M, 2^N ]
    iutil <- apply(iland, 1, function(x){
      tmpmat[i,] <- x  ## set counterfactual actor-component configuration
      config_util <- if('matrix' %in% class(theta)) {
        self$get_struct_mod_stats_mat_from_bi_mat( tmpmat ) %*% theta[step_id,] ## variable thetas
      } else {
        self$get_struct_mod_stats_mat_from_bi_mat( tmpmat ) %*% theta   ## fixed thetas
      }
      return( config_util ) ## compute utility vector
    }) 
    
   
    # plot(ifit[actor_i, ])
    
    # apply
    
    # ids.max <- which(ifit == max(ifit), arr.ind = TRUE) 
    # fits.max <- ifit[ ids.max ]
    # nmax <- length(fits.max)
    
    # ifit
    
    # ids.max[1,]
    
    # act_counterfacts[[i]] <- list(
    #   ifit = ifit, #Given all other ties, Actor i's configurations applied to utility func for all actors
    #   ids.max = ids.max,
    #   fits.max = fits.max,
    #   nmax = nmax
    # )
    
    ## distances of each counterfactual configuration 
    actor_i <- i
    start_actor_config <- bi_env_mat_step[actor_i, ]
    start_actor_config_str <- paste(start_actor_config, collapse = '|')
    dist_counterfac <- iland - matrix( rep(start_actor_config, 2^self$N), byrow=T, ncol=self$N)
    z <- t(sapply( 1:nrow(dist_counterfac),  function(config_i) {
        x <- dist_counterfac[config_i,  ]   ## dist_counterfac[MxN]
        util <- iutil[actor_i , config_i]  ## iutil[MxN]
        fit <- fit_vec[config_i]
        c(dist=sum(x!=0),
          drop=sum(x==-1), 
          nochange=sum(x==0),
          add=sum(x==1), 
          config_str = paste( iland[config_i, ], collapse = '|'),
          config_dist_str=paste(x, collapse = '|'), 
          fitness=fit, 
          utility=util
          )
      })) %>% tidyr::as_tibble() %>% 
      mutate(start_str=start_actor_config_str, 
             chain_step_id=step_id,
             fitness = as.numeric(fitness),
             utility = as.numeric(utility),
             dist = as.numeric(dist),
             drop = as.numeric(drop),
             nochange = as.numeric(nochange),
             add = as.numeric(add)
             )

plt <- z %>% ggplot(aes(x=utility, y=fitness, color=factor(add))) +
  geom_smooth(alpha=.1, method='lm') +
  geom_point(alpha=.2) + 
  facet_grid(factor(dist) ~ . , scales='free_y') +
  geom_vline(aes(xintercept = mean, color=factor(dist)), data=z%>%group_by(dist,chain_step_id)%>%summarize(mean=mean(utility))) +
# facet_grid(.~dist) +
  theme_bw() + 
  ggtitle(sprintf('Actor %s Fitness as function of Utility\nby configuration distance (number of tie changes from current config.)', actor_i))

print(plt)


(fita <- lm(fitness ~ utility*factor(add), data=z ))
print(summary(fita))
(fitd <- lm(fitness ~ utility*factor(drop), data=z ))
print(summary(fitd))
    
```


```{r, fig.height=8, fig.width=10, fig.dpi=800}
    self <- env1

theta <- self$rsiena_model$thetaUsed

# step_ids <- round(c(1,8,64,512, self$rsiena_model$n3/2, self$rsiena_model$n3))
acpds <- seq(1, nrow(theta), by= self$M )
step_ids <- c(acpds[1:min(3, length(acpds))])
#
step_id <- step_ids[ 2 ]

  # step_id <- 1
  cat(sprintf('\nstep %s: actors ', step_id))
  
  bi_env_mat_step <- self$bi_env_arr[, , step_id]
  
  act_counterfacts <- list()
  
  i <- 5
  
  # for (i in 1:min(3, self$M)) {
    cat(sprintf(' %s ', i))
    
    
    ##**ACTOR i DECISION PERSPECTIVE** 0000000000000000000000000000000000
    
    ##**TODO**
    ## ALL actor-component counterfactual configuations for Actor i  (2^N rows by N cols)
    iland <- expand.grid(lapply(1:self$N, function(x) 0:1 ))
    iland_config_step_row_id <- which(apply(iland, 1, function(x) all(x == bi_env_mat_step[i,]) ))
    
    fit_vec <- env1$fitness_landscape[1, , 1 + 2*env1$N]
    
    ## fitness
    fitness_land1 <- env1$fitness_landscape[1, , 1:env1$N]
    # #
    # if ( ! all( iland == fitness_land1) ) {
    #   stop('fitness landscape configurations do not match iland configurations')
    # }
    
    
    
    ##**TODO**
    ##**TODO** COMBINE utility and 'fitness' measures in dataframe by configuration with distance from actor's current configuration
    ##*##**TODO**
    ##**TODO**
    ##**TODO  CHANGE TO ACTOR-SPECIFIC LANDSCAPES **
    tmpmat <- bi_env_mat_step
    ## ifit dimensions [ M, 2^N ]
    iutil <- apply(iland, 1, function(x){
      tmpmat[i,] <- x  ## set counterfactual actor-component configuration
      config_util <- if('matrix' %in% class(theta)) {
        self$get_struct_mod_stats_mat_from_bi_mat( tmpmat ) %*% theta[step_id,] ## variable thetas
      } else {
        self$get_struct_mod_stats_mat_from_bi_mat( tmpmat ) %*% theta   ## fixed thetas
      }
      return( config_util ) ## compute utility vector
    }) 
    
   
    # plot(ifit[actor_i, ])
    
    # apply
    
    # ids.max <- which(ifit == max(ifit), arr.ind = TRUE) 
    # fits.max <- ifit[ ids.max ]
    # nmax <- length(fits.max)
    
    # ifit
    
    # ids.max[1,]
    
    # act_counterfacts[[i]] <- list(
    #   ifit = ifit, #Given all other ties, Actor i's configurations applied to utility func for all actors
    #   ids.max = ids.max,
    #   fits.max = fits.max,
    #   nmax = nmax
    # )
    
    ## distances of each counterfactual configuration 
    actor_i <- i
    start_actor_config <- bi_env_mat_step[actor_i, ]
    start_actor_config_str <- paste(start_actor_config, collapse = '|')
    dist_counterfac <- iland - matrix( rep(start_actor_config, 2^self$N), byrow=T, ncol=self$N)
    z <- t(sapply( 1:nrow(dist_counterfac),  function(config_i) {
        x <- dist_counterfac[config_i,  ]   ## dist_counterfac[MxN]
        util <- iutil[actor_i , config_i]  ## iutil[MxN]
        fit <- fit_vec[config_i]
        c(dist=sum(x!=0),
          drop=sum(x==-1), 
          nochange=sum(x==0),
          add=sum(x==1), 
          config_str = paste( iland[config_i, ], collapse = '|'),
          config_dist_str=paste(x, collapse = '|'), 
          fitness=fit, 
          utility=util
          )
      })) %>% tidyr::as_tibble() %>% 
      mutate(start_str=start_actor_config_str, 
             chain_step_id=step_id,
             fitness = as.numeric(fitness),
             utility = as.numeric(utility),
             dist = as.numeric(dist),
             drop = as.numeric(drop),
             nochange = as.numeric(nochange),
             add = as.numeric(add)
             )

plt <- z %>% ggplot(aes(x=utility, y=fitness, color=factor(drop))) +
  geom_smooth(alpha=.1, method='lm') +
  geom_point(alpha=.2) + 
  facet_grid(factor(dist) ~ . , scales='free_y') +
  geom_vline(aes(xintercept = mean, color=factor(dist)), data=z%>%group_by(dist,chain_step_id)%>%summarize(mean=mean(utility))) +
# facet_grid(.~dist) +
  theme_bw() + 
  ggtitle(sprintf('Actor %s Fitness as function of Utility\nby configuration distance (number of tie changes from current config.)', actor_i))

print(plt)


(fita <- lm(fitness ~ utility*factor(add), data=z ))
print(summary(fita))
(fitd <- lm(fitness ~ utility*factor(drop), data=z ))
print(summary(fitd))
    
```
    
    
    
```{r, include=FALSE}
    
    # # z <- apply( dist_counterfac, 1, function(x) {
    # #   c(dist=sum(x!=0),
    # #     `drop`=sum(x==-1), 
    # #     `nochange`=sum(x==0),
    # #     `add`=sum(x==1), 
    # #     counterfac=paste(x, collapse = '|'), 
    # #     start=paste(bi_env_mat_step[i, ], collapse = '|') 
    # #     )
    # # })
    # # z <- z[,order(z['d',], decreasing = F)]
    # ##**TODO**
    # ##**INFORMATION LOSS**
    # ##  This uses only ego's own counterfactual fits (landscape)
    # ##  but the corresponding other actor's affected fits (landscapes are not currently used )
    # # z <- rbind(z, utility_ego=ifit[ i , ] )
    # # z <- rbind(z, utility_alter_mean= colMeans(ifit[ -i , ], na.rm = T) )
    # # z <- rbind(z, utility_alter_sd  = apply(ifit[ -i , ], 2, function(x) sd(x, na.rm = T)) )
    # 
    # wdf <- as.data.frame( t(z) )
    # 
    # actfit_long <- wdf %>% 
    #   mutate(
    #     dist = as.numeric(dist),
    #     drop = as.numeric(drop),
    #     nochange = as.numeric(nochange),
    #     add = as.numeric(add) ,
    #     utility_ego=as.numeric(utility_ego),
    #     utility_alter_mean = as.numeric(utility_alter_mean),
    #     utility_alter_sd = as.numeric(utility_alter_sd)
    #   ) %>%
    #   pivot_longer(cols = c( drop:add, utility_ego:utility_alter_sd )) 
    #  
    # 
    # plt <- actfit_long %>% ggplot(aes(x=value))+ 
    #   # geom_density() + 
    #   geom_histogram() +
    #   facet_grid(dist ~ name, scales='free_x') + theme_bw() + 
    #   ggtitle(sprintf('Utility Transition Paths: Actor %s, Step %s', i, step_id))
    # 
    # pltlist[[ length(pltlist)+1 ]] <- plt
    # 
    # # z <- apply(iland, 1, function(x) {
    # #   x <- plyr::count( x - bi_env_mat_step[i, ]) 
    # #   cnts <- merge(x$freq, data.frame(x=c())
    # #   names(cnts) <- x$x
    # #   return(cnts)
    # # }, simplify = T)
    # # 
    # # stringdist::stringdist( , , method='hamming')
    # 
    # step_actor_key <- sprintf('%s|%s',step_id,i)
    # 
    # statsl[[step_actor_key]] <- list(
    #   actfit_long = actfit_long,
    #   act_counterfacts = act_counterfacts,
    #   plt = plt,
    #   actor_id=i, chain_step_id = step_id,
    #   strategy=as.factor(self$strat_1_coCovar[i])
    # )
    # 
    # actor_strats <- env3$get_actor_strategies()
    # 
    # utilist[[step_actor_key]] <- actfit_long %>% 
    #   filter(name %in% c('utility_ego','utility_alter_mean')) %>%
    #   mutate(actor_id=i, chain_step_id = step_id, strategy=as.factor(actor_strats[i]))
      
```


```{r}
heatmap(env1$fitness_landscape[5,,], col=heat.colors(50), scale = 'row')
heatmap(env1$fitness_landscape[6,,], col=heat.colors(50), scale = 'row')
```

```{r}
image(env1$fitness_landscape[2,,])

image(t(apply(env1$fitness_landscape[2,,], 2, rev)), col=heat.colors(50), axes=FALSE)

heatmap(env1$fitness_landscape[,,dim(env1$fitness_landscape)[3]-1], col=heat.colors(50), scale = 'row')
```


```{r}
ggheatmap::ggheatmap(env1$fitness_landscape[8,,])
ggheatmap::ggheatmap(env1$fitness_landscape[9,,])
ggheatmap::ggheatmap(env1$fitness_landscape[10,,])
```








```{r}
z$
```
