---
title: 'SaoMNK Tutorial: Sequential Exogenous Shock Comparison'
author: "S. Downing"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document:
    df_print: paged
  word_document: default
---

This project develops the idea that the task of endogenizing the 
NK search landscape can be accomplished by representing the search environment 
as a binary network of `M` actors affiliating with `N` components. 
The resulting bipartite network data generation process (DGP) -- 
using the bipartite network as the dependent variable (DV) -- 
enables dynamic network analysis of the search system as a form of 
Stochastic Actor-Oriented Model (SAOM) [(Snijders, 1996)](https://doi.org/10.1080/0022250X.1996.9990178).
This new Stochastic Actor-Oriented MNK (`SaoMNK`) Model reveals that the conventional 
`K` term (i.e., complexity, or degree of epistatic interdependence) is just 
one part (denoted `K_C` below) of a coupled distribution of network 
***degree*** terms (i.e., number of each type of tie to each type of node) 
between the search system's network nodes (Actors and/or Components):

 - ***2-mode***: Bipartite DGP
    1. `K_B1` = Actor **scope** (Actor-Component degree) 
    2. `K_B2` = Component **popularity** (Component-Actor degree)
 - ***1-mode***: Unipartite Projection 
    3. `K_A` = Actor **sociality** (Actor-[Component]-Actor degree)
    4. `K_C` = Component **complexity** *(Traditional `K`)* (Component-[Actor]-Component degree)

`SaoMNK` is designed for running strategic search simulation, testing, and 
experimentation by leveraging the `RSiena` package, an R implementation of
Simulation Investigation for Empirical Network Analysis (SIENA).

```{r dependencies, include=FALSE}
##** THIS BLOCK IS NOT SHOWN IN OUTPUT FILE (PDF,DOC,HTML) **##
rm(list=ls())
####
# ## General settings to hide code
# knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
knitr::opts_chunk$set(warning = FALSE)

# rm(list=ls())  ## uncomment to clear environ before run
########################
##
##   SAOM-NK Runscript:
##
##   1. Estimation Recovery 
##
##
#######################
library(plyr)
library(dplyr)
library(uuid)
library(ggpubr)
library(data.table)
library(DT)
library(Matrix)# For block diagonal matrices
library(knitr)
library(kableExtra)

#
.ts <- gsub('\\.','',as.character(as.numeric(Sys.time())))
#
# .uuid <- UUIDgenerate()
##  DIRECTORY
# grid.dir <- sprintf('_SAOMNK_sensitivity_start0_rand20__%s', .ts )
###############  Load R DEPENDENCIES ############################
## Directories
dir_ext <- 'D:\\Search_networks'
dir_r <- 'C:\\Users\\sdr8y\\OneDrive - University of Missouri\\Research\\Search_networks\\SaoMNK\\R'
###############  Load R6 Class DEPENDENCIES ############################
## Biparite Environment Search Simulation Class
SaoMNK <- source(file.path(dir_r, 'SAOM_NK_R6_model.R'))$value


#
create_block_diag <- function(N, B) {
  # Determine the approximate block size
  block_sizes <- rep(N %/% B, B)  # Equal-sized blocks
  block_sizes[1:(N %% B)] <- if (N %% B == 0) {
      block_sizes[1:(N %% B)]   # Adjust for remainder
    } else {
      block_sizes[1:(N %% B)] + 1
    }
  # Create individual binary blocks
  blocks <- lapply(block_sizes, function(s) matrix(1, nrow = s, ncol = s))
  # Combine blocks into a block diagonal matrix
  block_diag_matrix <- as.matrix(bdiag(blocks))  # Convert sparse to standard matrix
  # Ensure exact NxN dimensions
  block_diag_matrix <- block_diag_matrix[1:N, 1:N]  # Trim to N x N
  # return
  return(block_diag_matrix)
}
# # Example usage:
# N <- 6  # Number of rows/columns
# B <- 2  # Number of blocks
# block_matrix <- create_block_diag(8, 2)
# # Print the matrix
# print(block_matrix)

## default settings: Users do not change; TODO: implment within restricted class attributes
DV_NAME <- 'self$bipartite_rsienaDV'

```


```{r, warning=FALSE}
## Load SaoMNK Biparite Environment Search Simulation Class
SaoMNK <- source(file.path(dir_r, 'SAOM_NK_R6_model.R'))$value
## define simulation iterations based on 
steps_per_actor <- 36
```

# 1. Shock Sequence A:  (i)Short/Large (ii)Long/Small


## 1.1 Environment Configuration

```{r}
#
environ_params <- list(
  M = 9,       ## Actors
  N = 12,       ## Components
  BI_PROB = 0, ## Environmental Density (DGP hyperparameter)
  component_matrix_start = 'rand', ##**TODO** Implement: 'rand','modular','semi-modular',...
  rand_seed = 1234,
  plot_init = F,
  name = '_test_tutorial_nb_'
)
#
env1 <- SaoMNK$new(environ_params)
```



## 2.2  Structure Model


### 2.2.1 Component Interaction (Epistasis) Matrix

SAOM Objective Function serves as the stochastic actor's utility function for strategic search.

```{r}
#
strategies <- list(
  egoX   =   c(-1, 0, 1),
  inPopX =   c( 1, 0, -1)
)

## 2.b. Component Payoffs vector
set.seed(12345)
component_payoffs <-  runif(environ_params$N, min = 0, max = 1)
## 2. Strategies sets the objective function as a linear combination of network stats across DVs
#
actor_strats_list <- lapply(strategies, function(strat) rep(strat,  environ_params$M/length(strat)) )

component_int_mat <- create_block_diag(environ_params$N, 3)
component_int_mat
```


### 2.2.2 Structure Model List: RSiena parameters and covariates

```{r}
dyad_cov_X <- matrix(runif(n = environ_params$N*environ_params$M, min = -.5, max= 1), nrow=environ_params$M)
#
structure_model <- list(
  dv_bipartite = list(
    name = 'self$bipartite_rsienaDV',
    rates = list(
      # list(effect='Rate', parameter= .111, dv_name=DV_NAME, fix=FALSE )#, ##interaction1 = NULL
    ),
    effects = list( ##**STRUCTURAL EFFECTS -- dyadic/network endogeneity sources**
      list(effect='density', parameter= -1, dv_name=DV_NAME, fix=T ), ##interaction1 = NULL
      list(effect='outAct',  parameter= 0, dv_name=DV_NAME, fix=T ),
      list(effect='inPop',   parameter= 0, dv_name=DV_NAME, fix=T ) #interaction1 = NUL
    ),
    ## COVARIATE EFFECTS
    coCovars = list( 
      ##** COMPONENTS : MONADIC CONSTANT COVARIATE EFFECTS **##
      list(effect='altX',   parameter= 0, dv_name=DV_NAME, fix=T,
           interaction1='self$component_1_coCovar', x = component_payoffs
      ),
      # list(effect='outActX',   parameter= 0, dv_name=DV_NAME, fix=T,
      #      interaction1='self$component_1_coCovar', x = component_payoffs 
      # ),
      ##** STRATEGIES : MONADIC CONSTANT COVARIATE EFFECTS **##
      list(effect='egoX',   parameter= 0,  dv_name=DV_NAME, fix=T,
           interaction1='self$strat_1_coCovar',   x = actor_strats_list[[1]] 
      ), #interaction1 = NULL
      # list(effect='inPopX', parameter= 0,  dv_name=DV_NAME, fix=T,
      #      interaction1='self$strat_2_coCovar',  x = actor_strats_list[[2]]
      # ),
      list(effect='egoX', parameter= 0,  dv_name=DV_NAME, fix=T,
           interaction1='self$strat_2_coCovar',  x = actor_strats_list[[2]]
      )
    ),
    ##**MONADIC TIME-VARYING COVARIATE EFFECTS -- DYNAMIC STRATEGY PROGRAMS**
    varCovars = list(),
    ##**DYADIC CONSTANT COVARIATE EFFECTS -- EXOGENOUS INTERACTION MATRIX**
    coDyadCovars = list(
      list(effect='XWX',   parameter= .1, dv_name=DV_NAME, fix=T, nodeSet=c('COMPONENTS','COMPONENTS'),##if M=N, must provide nodeSet
           interaction1='self$component_1_coDyadCovar',
           x = component_int_mat ## component-[actor]-component dyads
      )#,
      # list(effect='XWX',   parameter= -.02, dv_name=DV_NAME, fix=T, nodeSet=c('COMPONENTS','COMPONENTS'), ##if M=N, must provide nodeSet
      #      interaction1='self$component_2_coDyadCovar',
      #      x = (1-component_int_mat) ## deltas = changes of payoff contributions from each actor-component
      # )
    ),
     ##**DYADIC TIME-VARYING COVARIATE EFFECTS -- DYNAMIC INTERACTION MATRIX**
    varDyadCovars = list(),
    interactions = list(
      # list(effect='egoX|XWX',   parameter= 0, dv_name=DV_NAME, fix=T,
      #      interaction1='self$strat_1_coCovar',
      #      interaction2='self$component_1_coDyadCovar'
      # ),
      # list(effect='inPopX|altX',   parameter= 0, dv_name=DV_NAME, fix=T,
      #      interaction1='self$strat_2_coCovar',
      #      interaction2='self$component_1_coCovar'
      # ),
      # list(effect='egoX|altX',   parameter= 0, dv_name=DV_NAME, fix=T,
      #      interaction1='self$strat_2_coCovar',
      #      interaction2='self$component_1_coCovar'
      # )
    )
  )
)
```





### 2.2.3 Preview Structure Model Effects

```{r}
env1$preview_effects(structure_model, filter=FALSE)
```

```{r}
# ## Uncomment if HTML output for interactive filterable data table
# env1$preview_effects(structure_model, filter=TRUE)
```





## 1.3 Run search with exogneous shocks to parameters

```{r}
theta_shocks <- list(
  list(effect_level=c('egoX_1','egoX_2'), parameter=c(0, 0),
       portion = 2, shock_on=0
  ),
  list(effect_level=c('egoX_1','egoX_2'), parameter=c(0, .8),
       portion = 1, shock_on=1, label='x2 Short/Lg.'
  ),
  list(effect_level=c('egoX_1','egoX_2'), parameter=c(.6, 0),
       portion = 2, shock_on=1, label='x1 Long/Sm.'
  ),
  list(effect_level=c('egoX_1','egoX_2'), parameter=c(0, 0),
       portion = 1, shock_on=0
  )
)

## Run Rsiena search using variable parameters in theta_matrix
env1$search_rsiena(
  structure_model = structure_model,
  iterations_per_actor = steps_per_actor, ## alternative to "iterations=500"
  theta_shocks = theta_shocks,
  digits = 4,
  run_seed = 12345
)
```



```{r}
# shocks <- env1$fit_rsiena_shocks(n_obs=8)
```


```{r, fig.height=6, fig.width=7, fig.dpi=800}
# screenreg(lapply(shocks, function(x)x$rsiena_model), single.row = T)
```



## 1.4 Print Bipartite Search Network Snapshots


```{r}
print(env1$component_1_coDyadCovar )
```

Snapshopts of the the biparite network and the social and component epistasis interactions are taken at fixed step intervals to show the evolving multidimensional coupled search environment (actors and components)

```{r, fig.height=5, fig.width=11}
snapshot_ids <- c( round(seq(1,  round(dim(env1$bi_env_arr)[3]*0.6), length.out=12) ), dim(env1$bi_env_arr)[3]  )
env1$plot_snapshots(snapshot_ids)
```

## 1.5 Plot Degrees 

Time series is simulated decision steps.

```{r, fig.height=10, fig.width=10, fig.dpi=900}
##
env1$plot_degree_4panel(loess_span = 0.25)
```


## 1.8 Plot Utility by Actor over Time (Simulated Decision Steps)

```{r, warning=FALSE, echo=FALSE}

# env1$plot_actor_utility() + 
#   ggtitle('Utility by Actor\nLoess Curves vs. Periodwise Avg. in Black')

```

## 1.9 Plot Contributions to Actor Utility colored by Strategy

Time series is simulated decision steps.

```{r,  fig.height=13, fig.width=9, fig.dpi=900}
##
env1$plot_utility_contributions(loess_span=0.25)
```

## 1.10 Plot Actor Utility Timeseries and Distribution by Strategy

```{r, fig.height=6, fig.width=8, fig.dpi=800}
env1$plot_actor_utility_strategy_summary(loess_span=0.3)
```

## 1.11 Fitness Landscape Peaks Distributions

Number of changes to component ties as distance from counterfactual affiliation configurations

```{r}
## Use random uniform values for all combinations like traditional NK model
# env1$compute_fitness_landscape(component_coCovar=NA)
```



## 1.12 Plot Exogenous Shock Descriptive Comparison

```{r, fig.height=10, fig.width=8, fig.dpi=800}
env1$plot_shocks()
```



```{r}
# ## Define function
# print_kable <- function(test_list, index) {
#   item <- test_list[[ index ]]
#   item_label <- names(test_list)[index]
#   caption <- sprintf("%s",gsub('__',' vs. ',item_label))
#   return( kable( item, digits=3, caption=caption) )
# }
```



## 1.13 Multiperiod Diff-in-Diff Test Comparison

```{r, warning=FALSE}
suppressMessages({ shock_res1 <- env1$test_shocks('did') })
```


```{r, warning=FALSE, fig.height=11, fig.width=10, fig.dpi=800}
shock_res1$plot 
```


















# 2. Shock Sequence B:  (i)Long/Small, (ii)Short/Large 




```{r}
env2 <- SaoMNK$new(environ_params)
```




## 2.3 Run RSiena Search Process

```{r}
theta_shocks_2 <- list(
  list(effect_level=c('egoX_1','egoX_2'), parameter=c(0, 0),
       portion = 2, shock_on=0
  ),
  list(effect_level=c('egoX_1','egoX_2'), parameter=c(.6, 0),
       portion = 2, shock_on=1, label='x1 Long/Sm.'
  ),
  list(effect_level=c('egoX_1','egoX_2'), parameter=c(0, .8),
       portion = 1, shock_on=1, label='x2 Short/Lg.'
  ),
  list(effect_level=c('egoX_1','egoX_2'), parameter=c(0, 0),
       portion = 1, shock_on=0
  )
)


## Run Rsiena search using variable parameters in theta_matrix
env2$search_rsiena(
  structure_model = structure_model,
  theta_shocks = theta_shocks_2,
  iterations_per_actor = steps_per_actor,
  digits = 4,
  run_seed = 12345
)
```


```{r}
# shocks2 <- env2$fit_rsiena_shocks(n_obs=10)

```


```{r, fig.height=6, fig.width=7, fig.dpi=800}
# screenreg(lapply(shocks2, function(x)x$rsiena_model), single.row = T)
  
```

```{r}
## 1st and last state of the bipartite matrix system
print(env2$bipartite_matrix_init )
print(env2$bi_env_arr[,, dim(env2$bi_env_arr)[3] ] )
```

```{r}
print(env2$component_1_coDyadCovar )
```

## 2.4 Print Bipartite Search Network Snapshots

Snapshopts of the the biparite network and the social and component epistasis interactions are taken at fixed step intervals to show the evolving multidimensional coupled search environment (actors and components)

```{r, fig.height=5, fig.width=11}
snapshot_ids <- c( round(seq(1,  round(dim(env2$bi_env_arr)[3]*0.6), length.out=12) ), dim(env2$bi_env_arr)[3]  )
env2$plot_snapshots(snapshot_ids)
```

## 2.5 Plot Degrees 

Time series is simulated decision steps.

```{r, fig.height=10, fig.width=10, fig.dpi=900}
##
env2$plot_degree_4panel(loess_span = 0.25)
```


## 2.8 Plot Utility by Actor over Time (Simulated Decision Steps)

```{r, warning=FALSE, echo=FALSE}

# env2$plot_actor_utility() + 
#   ggtitle('Utility by Actor\nLoess Curves vs. Periodwise Avg. in Black')

```

## 2.9 Plot Contributions to Actor Utility colored by Strategy

Time series is simulated decision steps.

```{r,  fig.height=13, fig.width=9, fig.dpi=900}
##
env2$plot_utility_contributions(loess_span=0.25)
```

## 2.10 Plot Actor Utility Timeseries and Distribution by Strategy

```{r, fig.height=6, fig.width=8, fig.dpi=800}
env2$plot_actor_utility_strategy_summary(loess_span=0.3)
```

## 2.11 Fitness Landscape Peaks Distributions

Number of changes to component ties as distance from counterfactual affiliation configurations

```{r}
## Use random uniform values for all combinations like traditional NK model
# env2$compute_fitness_landscape(component_coCovar=NA)
```




## 2.12 Plot Exogenous Shock Descriptive Comparison

```{r, fig.height=10, fig.width=8, fig.dpi=800}
env2$plot_shocks()
```





## 2.13 Multiperiod Diff-in-Diff Test Comparison

```{r, warning=FALSE}
suppressMessages({ shock_res2 <- env2$test_shocks('did') })
```


```{r, warning=FALSE, fig.height=11, fig.width=10, fig.dpi=800}
shock_res2$plot 
```


















