---
title: 'SaoMNK Tutorial: Exogenous Shocks'
author: "S. Downing"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

This project begins from the simple premise that the task of endogenizing the NK search landscape can be done by representing the search environment as a binary bipartite network of `M` actors affiliating with `N` components. This bipartite network can be then be analyzed according to the Stochastic Actor-Oriented Model (SAOM) [(Snijders, 1996)](https://doi.org/10.1080/0022250X.1996.9990178).

Thus, the Stochastic Actor-Oriented MNK model, abbrivated `SaoMNK`, is designed for running strategic search simulation, testing, and experimentation by leveraging the `RSiena` package, an R implementation of\
Simulation Investigation for Empirical Network Analysis (SIENA).

This tutorial offers a basic introduction to `SaoMNK`.

```{r dependencies, include=FALSE}
rm(list=ls())
####
# ## General settings to hide code
# knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
knitr::opts_chunk$set(warning = FALSE)

# rm(list=ls())  ## uncomment to clear environ before run
########################
##
##   SAOM-NK Runscript:
##
##   1. Estimation Recovery 
##
##
#######################
library(plyr)
library(dplyr)
library(uuid)
library(ggpubr)
library(data.table)
library(DT)
library(Matrix)  # For block diagonal matrices

#
.ts <- gsub('\\.','',as.character(as.numeric(Sys.time())))
#
# .uuid <- UUIDgenerate()
##  DIRECTORY
# grid.dir <- sprintf('_SAOMNK_sensitivity_start0_rand20__%s', .ts )
###############  Load R DEPENDENCIES ############################
## Directories
dir_ext <- 'D:\\Search_networks'
dir_r <- 'C:\\Users\\sdr8y\\OneDrive - University of Missouri\\Research\\Search_networks\\SaoMNK\\R'
###############  Load R6 Class DEPENDENCIES ############################
## Biparite Environment Search Simulation Class
SaomNkRSienaBiEnv <- source(file.path(dir_r, 'SAOM_NK_R6_model.R'))$value


#
create_block_diag <- function(N, B) {
  # Determine the approximate block size
  block_sizes <- rep(N %/% B, B)  # Equal-sized blocks
  block_sizes[1:(N %% B)] <- if (N %% B == 0) {
    block_sizes[1:(N %% B)]   # Adjust for remainder
  } else {
    block_sizes[1:(N %% B)] + 1
  }
  # Create individual binary blocks
  blocks <- lapply(block_sizes, function(s) matrix(1, nrow = s, ncol = s))
  # Combine blocks into a block diagonal matrix
  block_diag_matrix <- as.matrix(bdiag(blocks))  # Convert sparse to standard matrix
  # Ensure exact NxN dimensions
  block_diag_matrix <- block_diag_matrix[1:N, 1:N]  # Trim to N x N
  # return
  return(block_diag_matrix)
}
# # Example usage:
# N <- 6  # Number of rows/columns
# B <- 2  # Number of blocks
# block_matrix <- create_block_diag(8, 2)
# # Print the matrix
# print(block_matrix)

```

```{r, warning=FALSE}
###############  Load R6 Class DEPENDENCIES ############################
## Biparite Environment Search Simulation Class
SaomNkRSienaBiEnv <- source(file.path(dir_r, 'SAOM_NK_R6_model.R'))$value
## default settings: Users do not change; TODO: implment within restricted class attributes
DV_NAME <- 'self$bipartite_rsienaDV'
## define simulation iterations based on 
steps_per_actor <- 12
```

# 1. Baseline Case: No Shocks

## 1.1 Environment Configuration

```{r}
#
environ_params <- list(
  M = 12,       ## Actors
  N = 12,       ## Components
  BI_PROB = 0, ## Environmental Density (DGP hyperparameter)
  component_matrix_start = 'rand', ##**TODO** Implement: 'rand','modular','semi-modular',...
  rand_seed = 1234,
  plot_init = F,
  name = '_test_tutorial_nb_'
)
#
env1 <- SaomNkRSienaBiEnv$new(environ_params)
```


## 2.2 Structrual Model

SAOM Objective Function serves as the stochastic actor's utility function for strategic search.

```{r}
#
strategies <- list(
  egoX   =   c(-1, 0, 1),
  inPopX =   c( 1, 0, -1)
)

## 2.b. Component Payoffs vector
set.seed(12345)
component_payoffs <-  runif(environ_params$N, min = 0, max = 1)
## 2. Strategies sets the objective function as a linear combination of network stats across DVs
#
actor_strats_list <- lapply(strategies, function(strat) rep(strat,  environ_params$M/length(strat)) )

component_int_mat <- create_block_diag(environ_params$N, round(environ_params$N/4))

# dyad_cov_XWX <- ( outer(component_payoffs, component_payoffs, '*') * 
#     create_block_diag(environ_params$N, round(environ_params$N/3))
# )

dyad_cov_X <- matrix(runif(n = environ_params$N*environ_params$M, min = -.5, max= 1), nrow=environ_params$M)
# dyad_cov_X <- matrix(runif(environ_params$N*environ_params$M) - runif(environ_params$N*environ_params$M), 
#                      nrow=environ_params$M) 

# dyad_cov_XWX_X <- t( dyad_cov_XWX %*% t(dyad_cov_X) )

#
structure_model <- list(
  dv_bipartite = list(
    name = 'self$bipartite_rsienaDV',
    rates = list(
      list(effect='Rate', parameter= .111, dv_name=DV_NAME, fix=FALSE )#, ##interaction1 = NULL
    ),
    effects = list( ##**STRUCTURAL EFFECTS -- dyadic/network endogeneity sources**
      list(effect='density', parameter= 0, dv_name=DV_NAME, fix=T ), ##interaction1 = NULL
      list(effect='outAct',  parameter= 0, dv_name=DV_NAME, fix=T ),
      list(effect='inPop',   parameter= 0, dv_name=DV_NAME, fix=T ) #interaction1 = NUL
    ),
    ## COVARIATE EFFECTS
    coCovars = list( 
      ##** COMPONENTS : MONADIC CONSTANT COVARIATE EFFECTS **##
      list(effect='altX',   parameter= 1, dv_name=DV_NAME, fix=T,
           interaction1='self$component_1_coCovar', x = component_payoffs
      ),
      # list(effect='outActX',   parameter= 0, dv_name=DV_NAME, fix=T,
      #      interaction1='self$component_1_coCovar', x = component_payoffs 
      # ),
      ##** STRATEGIES : MONADIC CONSTANT COVARIATE EFFECTS **##
      list(effect='egoX',   parameter= 1,  dv_name=DV_NAME, fix=T,
           interaction1='self$strat_1_coCovar',   x = actor_strats_list[[1]] 
      ), #interaction1 = NULL
      list(effect='inPopX', parameter= 1,  dv_name=DV_NAME, fix=T,
           interaction1='self$strat_2_coCovar',  x = actor_strats_list[[2]]
      ) 
    ),
    ##**MONADIC TIME-VARYING COVARIATE EFFECTS -- DYNAMIC STRATEGY PROGRAMS**
    varCovars = list(),
    ##**DYADIC CONSTANT COVARIATE EFFECTS -- EXOGENOUS INTERACTION MATRIX**
    coDyadCovars = list(
      list(effect='XWX',   parameter= .1, dv_name=DV_NAME, fix=T, nodeSet=c('COMPONENTS','COMPONENTS'),##if M=N, must provide nodeSet
           interaction1='self$component_1_coDyadCovar',
           x = component_int_mat ## component-[actor]-component dyads
      ) #,
      # list(effect='X',   parameter= 0, dv_name=DV_NAME, fix=T, nodeSet=c('ACTORS','COMPONENTS'), ##if M=N, must provide nodeSet
      #      interaction1='self$component_2_coDyadCovar',
      #      x = dyad_cov_X ## deltas = changes of payoff contributions from each actor-component
      # )
    ),
     ##**DYADIC TIME-VARYING COVARIATE EFFECTS -- DYNAMIC INTERACTION MATRIX**
    varDyadCovars = list(),
    interactions = list(
      # list(effect='egoX|XWX',   parameter= 0, dv_name=DV_NAME, fix=T,
      #      interaction1='self$strat_1_coCovar',
      #      interaction2='self$component_1_coDyadCovar'
      # ),
      # list(effect='inPopX|altX',   parameter= 0, dv_name=DV_NAME, fix=T,
      #      interaction1='self$strat_2_coCovar',
      #      interaction2='self$component_1_coCovar'
      # ),
      # list(effect='egoX|altX',   parameter= 0, dv_name=DV_NAME, fix=T,
      #      interaction1='self$strat_2_coCovar',
      #      interaction2='self$component_1_coCovar'
      # )
    )
  )
)
```




```{r}
env1$preview_effects(structure_model, filter=FALSE)
```

```{r}
# ## Uncomment for HTML output with filterable data table
# env1$preview_effects(structure_model, filter=TRUE)
```





## 1.3 Run search with exogneous shocks to parameters

```{r}

theta_shocks <- list(
  ## only one period of shock is supported currently
  s01 = list(effect=c('egoX','inPopX'), parameter=c( 1, 1), portion = 1, shock_on=0),
  s11 = list(effect=c('egoX','inPopX'), parameter=c(-2,-2), portion = 1, shock_on=1, label='Short Shock'),
  s02 = list(effect=c('egoX','inPopX'), parameter=c( 1, 1), portion = 1, shock_on=0),
  s12 = list(effect=c('egoX','inPopX'), parameter=c(-2,-2), portion = 3, shock_on=1, label='Long Shock'),
  s03 = list(effect=c('egoX','inPopX'), parameter=c( 1, 1), portion = 2, shock_on=0)
)

## Run Rsiena search using variable parameters in theta_matrix
env1$search_rsiena(
  structure_model,
  iterations_per_actor = steps_per_actor, ## alternative to "iterations=500"
  theta_shocks = theta_shocks,
  digits = 4,
  run_seed = 12345
)
```


## 1.4 Print Bipartite Search Network Snapshots


```{r}
print(env1$component_1_coDyadCovar )
```

Snapshopts of the the biparite network and the social and component epistasis interactions are taken at fixed step intervals to show the evolving multidimensional coupled search environment (actors and components)

```{r, fig.height=5, fig.width=11}
snapshot_ids <- c(1, (1:8)*env1$M, dim(env1$bi_env_arr)[3]  )
env1$plot_snapshots(snapshot_ids)
```

## 1.5 Plot Actor Degrees (Component Scope and Common Affiliation Social Ties)

Time series is simulated decision steps.

```{r, fig.height=6, fig.width=6.5, fig.dpi=800}
##
env1$plot_actor_degrees(loess_span = 0.5)
```

## 1.6 Plot Component Degrees (Membership/Popularity and Common Actor Epistasis Ties)

```{r, fig.height=6, fig.width=6.5, fig.dpi=800}
##
env1$plot_component_degrees(loess_span = 0.5)

```

## 1.7 Plot Utility by Strategy over Time (Simulated Decision Steps)

```{r, warning=FALSE, echo=FALSE}

env1$plot_strategy_utility() + 
  ggtitle('Utility by Strategy\nLoess Curves vs. Periodwise Avg. in Black')

```

## 1.8 Plot Utility by Actor over Time (Simulated Decision Steps)

```{r, warning=FALSE, echo=FALSE}

# env1$plot_actor_utility() + 
#   ggtitle('Utility by Actor\nLoess Curves vs. Periodwise Avg. in Black')

```

## 1.9 Plot Contributions to Actor Utility colored by Strategy

Time series is simulated decision steps.

```{r,  fig.height=13, fig.width=9, fig.dpi=900}
##
env1$plot_utility_contributions(loess_span=0.25)
```

## 1.10 Plot Actor Utility Timeseries and Distribution by Strategy

```{r, fig.height=6, fig.width=8, fig.dpi=800}
env1$plot_actor_utility_strategy_summary(loess_span=0.5)
```

## 1.11 Fitness Landscape Peaks Distributions

Number of changes to component ties as distance from counterfactual affiliation configurations

```{r}
## Use random uniform values for all combinations like traditional NK model
env1$compute_fitness_landscape(component_coCovar=NA)
```





```{r}
env2 <- SaomNkRSienaBiEnv$new(environ_params)
```




## 2.3 Run RSiena Search Process

```{r}

## TODO:  PICK UP WITH coDydCovar Interation Matrix

## Run Rsiena search using variable parameters in theta_matrix
env2$search_rsiena(
  structure_model,
  iterations_per_actor = steps_per_actor,
  digits = 4,
  run_seed = 12345
)
```

```{r}

# saomResDf <- function(res, digits=3) 
# {
#   df <- data.frame(
#     DV=res$effects$name,
#     Effect=res$effects$effectName,
#     Type=res$effects$type,
#     Est=c( NA, round(res$theta, digits = digits)),
#     se=c(NA, round(res$se, digits = digits)),
#     t=c(NA, round(res$theta/res$se, digits = digits)),
#     p=c(NA, round(pt(abs(res$theta/res$se), df=Inf, lower.tail = F) * 2, digits = digits)),
#     stringsAsFactors = F
#   )
#   idx.rate <- grep('rate',df$Effect,T,T)
#   df$t[idx.rate] <- NA
#   df$p[idx.rate] <- NA
#   return(df)
# }
# 
# saomResDf(env2$rsiena_model)
```

```{r}

# ##
# #  Create SAOM regression comparison Table
# ##
# saomTable <- function(resList, file=NA, nameMap=NA, digits=3, print=TRUE,
#                       se.brackets=c('(',')'),p.brackets=c('[',']'),
#                       lin.comb.test.eval.endow=c('transitive triads'),
#                       show.sig.stars=TRUE, drop.p.col=TRUE, drop.dv.col=TRUE)
# {
#   if (se.brackets[1]=='(') se.brackets[1] <- '\'('
#   if ( p.brackets[1]=='(')  p.brackets[1] <- '\'(' 
#   
#   if (class(resList) != 'list') {
#     resList <- list(resList)
#   }
#   behNames <- c()
#   netNames <- c()
#   for (res in resList) {
#     behNames <- c(behNames, names(res$f$Data1$Behaviors))
#     netNames <- c(netNames, names(res$f$Data1$bipartites))
#     # netNames <- c(netNames, names(res$f$Data1$nets))
#   }
#   dvNames <- unique(c(behNames, netNames))
#   
#   nameList <- list()
#   for (dv in dvNames) {
#     resDvdf <- data.frame()
#     for (res in resList) {
#       .effDf <- as.data.frame(res$effects)
#       # effTypeNames <- .effDf[grep(dv, res$effects$name,T,T), c('effectName','type')]
#       effTypeNames <- .effDf[which(.effDf$name==dv & .effDf$include==T), c('effectName','type') ]
#       resDvdf <- unique(rbind(resDvdf, effTypeNames))
#     }
#     nameList[[dv]] <- unique(rbind(resDvdf, effTypeNames))
#   }
#   
#   dfl <- list()
#   for (res in resList) {
#     for (dv in dvNames) {
#       if (dv %in% res$effects$name) {
#         
#         # ## results dataframe combining both DVs
#         .df <- saomResDf(res, digits=digits)
#         
#         ##** add linear combination test parameter **
#         effdf <- res$requestedEffects
#         if( any(str_length(lin.comb.test.eval.endow) > 0) ) {
#           
#           id.eval  <- which( effdf$type == 'eval'  & effdf$effectName == lin.comb.test.eval.endow)
#           id.endow <- which( effdf$type == 'endow' & effdf$effectName == lin.comb.test.eval.endow)
#           ##** Recovering Maintenance Effect 2-sided p-val **
#           ##** based on covariance matrix rows for tested parameters **
#           #
#           lc.est <- .df$Est[id.eval] + .df$Est[id.endow]
#           #
#           lc.se <- c( sqrt(wt.mat %*% res$covtheta %*% t(wt.mat)) )[1]
#           #
#           lc.t <- lc.est / lc.se
#           lc.p <- pnorm(lc.t, lower.tail = T) * 2 ## two sided
#           #
#           ##** Wald Test Recovering Maintenance Effect **
#           # # ## Alternative Chi-sq (df=1) test to produce p-val 
#           # #    (same as two-sided p-val of norm dist with std.err. above)
#           # nparams <- nrow( res$requestedEffects )
#           # wt.mat <- matrix(0, 1, nparams)
#           # wt.mat[1,id.eval] <- 1
#           # wt.mat[1,id.endow] <- 1
#           # wt.out <- Wald.RSiena(wt.mat, res)
#           # lc.chisq <- wt.out$chisquare
#           # lc.df <- wt.out$df
#           # lc.1side.z <- c(wt.out$onesided)[1]
#           # lc.2side.p <-  c(wt.out$pvalue)
#         }
#         ## Append linear combination test to effects dataframe
#         .tmp <- data.frame(
#           DV = dv,
#           Effect = lin.comb.test.eval.endow,
#           Type = 'linear_combination',
#           Est = lc.est,
#           se = lc.se,
#           t = lc.t,
#           p = lc.p
#         )
#         ## append to top if linear combination is in first DV; else append to bottom
#         .df <- if (.df$DV[1]==dv){ rbind(.tmp, .df) }else{ rbind(.df, .tmp) }
#         ## append to output list
#         dfl[[dv]][[ length(dfl[[dv]]) + 1 ]] <- .df[which(.df$DV==dv),]
#       }
#     }
#   }
#   
#   mod.cols <- c('Est','se','p')
#   
#   tdf <- data.frame(stringsAsFactors = F)
#   for (dv in names(nameList)) {
#     hasRowname <- FALSE
#     for (rowi in 1:nrow(nameList[[dv]])) { ## effect row
#       eff <- nameList[[dv]][rowi,]
#       effRow <- list()
#       for (modDf in dfl[[dv]] ) { ## model dataframe in DV group
#         effId <- which(modDf$Effect == eff$effectName & modDf$Type == eff$type)
#         if (length(effId) > 0) {
#           effRow[[length(effRow)+1]] <- modDf[effId,mod.cols]
#         } else {
#           .nadf <- data.frame()
#           for (col in mod.cols) .nadf[1,col] <- NA
#           effRow[[length(effRow)+1]] <- .nadf
#         }
#       }
#       effRowDf <- cbind(data.frame(DV=dv,Effect=eff$effectName, Type=eff$type, stringsAsFactors = F), cbindDfList(effRow))
#       if (!hasRowname) {
#         effRowDf <- rbind(effRowDf, effRowDf)
#         effRowDf[1,]  <- c('', sprintf('Dynamics: %s', dv), rep(NA, ncol(effRowDf)-2))
#         hasRowname <- TRUE
#       }
#       tdf <- rbind(tdf, effRowDf)
#     }
#   }
#   
#   # move rate rows to end
#   .tmp.rate.row <- tdf[1,]
#   .tmp.rate.row$Effect <- 'Rate Parameters'
#   rate.idx <- which(tdf$Type=='rate')
#   tdf <- rbind(tdf[-rate.idx,], .tmp.rate.row, tdf[rate.idx, ])
# 
#   
#   ##
#   obs <- c()
#   ns <- c()
#   conv <- c()
#   convt <- c()
#   iter <- c()
#   for (res in resList) {
#     obs <-c(obs, res$observations)
#     ns <- c(ns, attributes(res$f$Data1$nets[[1]][[1]][[1]])$nActors)
#     conv <- c(conv, res$tconv.max)
#     convt<- c(convt, max(abs(res$tconv)))
#     iter <- c(iter, res$n)
#   }  
#   
#   # est idx
#   idx.est <- which(names(tdf)%in% 'Est')
#   #
#   tdf[nrow(tdf)+1, ] <- NA
#   tdf[nrow(tdf), 'Effect'] <- 'Time Periods'
#   tdf[nrow(tdf), idx.est] <-  obs
#   #
#   tdf[nrow(tdf)+1, ] <- NA
#   tdf[nrow(tdf), 'Effect'] <- 'Num. Firms'
#   tdf[nrow(tdf), idx.est] <-  ns
#   #
#   tdf[nrow(tdf)+1, ] <- NA
#   tdf[nrow(tdf), 'Effect'] <- 'Largest converg. t ratio'
#   tdf[nrow(tdf), idx.est] <-  round(convt, digits = digits)
#   #
#   tdf[nrow(tdf)+1, ] <- NA
#   tdf[nrow(tdf), 'Effect'] <- 'Overall max. converg. ratio'
#   tdf[nrow(tdf), idx.est] <-  round(conv, digits = digits)
#   #
#   tdf[nrow(tdf)+1, ] <- NA
#   tdf[nrow(tdf), 'Effect'] <- 'Iterations'
#   tdf[nrow(tdf), idx.est] <-  iter
#   
#   idx.se  <- grep('^se\\.{0,1}',names(tdf),T,T)
#   idx.p <- grep('^p\\.{0,}', names(tdf),T,T)
#   for (i in 1:length(idx.se)) {
#     sei <- idx.se[i]
#     pi <- idx.p[i]
#     ## Format Standard Error Column
#     tdf[,sei] <- apply(tdf[,c(sei,pi)],1,function(x) {
#       se <- x[1]
#       p <- x[2]
#       spfstr <- if(show.sig.stars) {
#         sprintf('%s%s%s.%sf%s%s',se.brackets[1],'%',digits+2,digits,se.brackets[2],getPvalStars(p))
#       } else {
#         sprintf('%s%s%s.%sf%s',se.brackets[1],'%',digits+2,digits,se.brackets[2])
#       }
#       ifelse(is.na(se)|se=='NA',' ',sprintf(spfstr,as.numeric(se)))
#     })
#     ## Format p-Value column
#     tdf[,pi] <- apply(tdf[,c(sei,pi)],1,function(x) {
#       se <- x[1]
#       p <- x[2]
#       pIsNum <-  !is.na(p) & p!='NA'
#       spfstr <- sprintf('%s%s%s.%sf%s',p.brackets[1],'%',digits+4,digits,p.brackets[2])
#       # cat(sprintf('\np = %s,  se = %s,  pIsNum = %s, p.spfstr = %s \n',p,se,pIsNum,spfstr))
#       ifelse(!pIsNum, ' ', 
#              ifelse(p<0.001,'[<0.001]',sprintf(spfstr, as.numeric(p))))
#     })
#   }
#   
#   ## add Type to name (not eval or rate)
#   idx.nonrate <- which(tdf$Type %in% c('endow','creation'))
#   tdf$Effect[idx.nonrate] <- apply(tdf[idx.nonrate,c('Effect','Type')],1,function(x){
#     sprintf('%s: %s',x[2],x[1])
#   })
#   tdf <- tdf[,which(names(tdf) != 'Type')]
#   
#   ## name mapping for effects
#   if (!any(is.na(nameMap))) 
#   {
#     ord <- c()
#     for (eff in names(nameMap)) {
#       idx.nm <- which(tdf$Effect == eff)
#       ord <- c(ord, idx.nm)
#       if (length(idx.nm) > 0) {
#         tdf$Effect[idx.nm] <- nameMap[[eff]]
#       }
#     }
#     tdf <- rbind(tdf[ord,], tdf[-ord,])
#   }
#   
#   if (drop.dv.col) {
#     tdf <- tdf[,-1]
#   }
#   if (drop.p.col) {
#     idx.p.col <- grep('^p[\\.\\d]{0,}',names(tdf),T,T)
#     tdf <- tdf[, -idx.p.col]
#   }
#   
#   if (print) {
#     print(tdf)
#   }
#   
#   if (!is.na(file)) {
#     write.csv(tdf, file = file, na = "", row.names = F)
#     cat(sprintf('\nFile written to %s\n', file))
#   }
#   
#   return(tdf)
#   
# }
# saomTable(list(env2$rsiena_model))
```

```{r}
# model <- env2$rsiena_model
# # Extract parameter estimates
# estimates <- model$theta
# 
# # Extract standard errors from the covariance matrix
# std_errors <- sqrt(diag(model$covtheta))
# 
# # Get effect names
# effect_names <- model$effects$effectName
# 
# # Compute t-ratios
# t_ratios <- estimates / std_errors
# 
# # Compute p-values (assuming normal distribution)
# p_values <- 2 * (1 - pnorm(abs(t_ratios)))
# 
# # Function to assign significance stars
# get_significance <- function(p_value) {
#   if (p_value < 0.001) return("***")
#   else if (p_value < 0.01) return("**")
#   else if (p_value < 0.05) return("*")
#   else return("")
# }
# 
# # Apply significance function
# significance_stars <- sapply(p_values, get_significance)
# 
# # Create a results data frame
# results_table <- data.frame(
#   Effect = effect_names,
#   Estimate = round(estimates, 3),
#   `Std. Error` = round(std_errors, 3),
#   `t-Ratio` = round(t_ratios, 3),
#   Significance = significance_stars
# )
# 
# # Print results nicely
# print(results_table)
# 
# # For better formatting in RMarkdown:
# library(knitr)
# kable(results_table, caption = "RSiena Model Results", digits = 3)
```

```{r}
## 1st and last state of the bipartite matrix system
print(env2$bipartite_matrix_init )
print(env2$bi_env_arr[,, dim(env2$bi_env_arr)[3] ] )
```

```{r}
print(env2$component_1_coDyadCovar )
```

## 2.4 Print Bipartite Search Network Snapshots

Snapshopts of the the biparite network and the social and component epistasis interactions are taken at fixed step intervals to show the evolving multidimensional coupled search environment (actors and components)

```{r, fig.height=5, fig.width=11}
snapshot_ids <- c(1, (1:8)*env2$M, dim(env2$bi_env_arr)[3]  )
env2$plot_snapshots(snapshot_ids)
```

## 2.5 Plot Actor Degrees (Component Scope and Common Affiliation Social Ties)

Time series is simulated decision steps.

```{r, fig.height=6, fig.width=6.5, fig.dpi=800}
##
env2$plot_actor_degrees(loess_span = 0.5)
```

## 2.6 Plot Component Degrees (Membership/Popularity and Common Actor Epistasis Ties)

```{r, fig.height=6, fig.width=6.5, fig.dpi=800}
##
env2$plot_component_degrees(loess_span = 0.5)

```

## 2.7 Plot Utility by Strategy over Time (Simulated Decision Steps)

```{r, warning=FALSE, echo=FALSE}

env2$plot_strategy_utility() + 
  ggtitle('Utility by Strategy\nLoess Curves vs. Periodwise Avg. in Black')

```

## 2.8 Plot Utility by Actor over Time (Simulated Decision Steps)

```{r, warning=FALSE, echo=FALSE}

# env2$plot_actor_utility() + 
#   ggtitle('Utility by Actor\nLoess Curves vs. Periodwise Avg. in Black')

```

## 2.9 Plot Contributions to Actor Utility colored by Strategy

Time series is simulated decision steps.

```{r,  fig.height=13, fig.width=9, fig.dpi=900}
##
env2$plot_utility_contributions(loess_span=0.25)
```

## 2.10 Plot Actor Utility Timeseries and Distribution by Strategy

```{r, fig.height=6, fig.width=8, fig.dpi=800}
env2$plot_actor_utility_strategy_summary(loess_span=0.5)
```

## 2.11 Fitness Landscape Peaks Distributions

Number of changes to component ties as distance from counterfactual affiliation configurations

```{r}
## Use random uniform values for all combinations like traditional NK model
env2$compute_fitness_landscape(component_coCovar=NA)
```



End.
